\documentclass[11pt]{letter}  %This tells LaTeX which type of document to use. I usually use letter for exams and worksheets, but there are also article, book, and ams art.
\usepackage{amsmath,amsthm} %Load the American Mathematical Society's theorem styles and mathematical formatting.
\usepackage{amsfonts} %Load the AMS fonts
\usepackage{amssymb} %Load the AMS symbols 
\usepackage{enumerate} %Load a package for prettier numbered lists
\usepackage[margin=1in]{geometry} %Set the margins
\usepackage{cite} %Load a package for bibliography, if we needed one
\usepackage{asymptote} %Load a package for drawing pictures

%Any of the packages can be removed, if you don't have them. The only essential packages are amsmath, amsthm,  amsfonts, and amssymb.


\begin{document} %The document is between the begin and end tags
\pagestyle{empty} %This stops LaTeX from numbering the pages



{\Large MA 225 Worksheet: Useful Logical Facts}\\ %\\ is a linebreak


The following logical facts can be deduced by looking at appropriate truth tables,  by formulating a tautologous propositional formula, or by just thinking it through.
\begin{center}
\begin{tabular}{c||c|c|c|c|c||}

$$ & 1 & 2 & 3 & 4 & 5 \\ \hline 
a & P & Q & $P \Rightarrow Q$ & P $\wedge$ Q & P $\vee$ Q \\ \hline
b & True & True & True & True & True \\ \hline
c & True & False & False & False & True \\ \hline 
d & False & True & True & False & True \\ \hline
e & False & False & True & False & False \\ \hline
\end{tabular}
\end{center}
\begin{enumerate}[(a)]
				\item If $P\Rightarrow Q$ is true and $P$ is true, then $Q$ must be true.\\
Proof:  Using the proof table above, we can use rows b and c which show when P is true. Then, looking on row b and c in columns 2 and 3, we can see that the only value which results in P $\Rightarrow$ Q being True is Q being True. \\
				\item If $P\Rightarrow Q$ is true and $Q$ is false, then $P$ must be false. \\
Proof: Looking at the proof table, we can see based off column 2 and 3, the only row which satisfies this is row e. \\
				\item If $P\wedge Q$ is true, then $P$ is true; if $P$ is true, then $P\vee Q$ is true. \\
Proof: We can prove the statement by looking at columns 1 2 and 4 of the truth table above to see that the only time P$\wedge$ Q is True is when P and Q are true, therefore P is True. Next, we can look at rows b and c where P are true and column 5 to see that when P is true, so is $P \vee Q$. \\
\begin{center}
\begin{tabular}{c||c|c|c|c|c|c||}

$$ & 1 & 2 & 3 & 4 & 5 & 6 \\ \hline
$$ & P & Q & R & P$ \Rightarrow Q$ & Q $\Rightarrow R$ R & $P \Rightarrow R$ \\ \hline
a & True & True & True & True & True & True \\ \hline
b & True & True & False & True & False & False \\ \hline
c & True & False & True & False & True & True \\ \hline
d & True & False & False & False & True & False \\ \hline
e & False & True & True & True & True & True \\ \hline
f & False & True & False & True & False & True \\ \hline
g & False & False & True & True & True & True \\ \hline
h & False & False & False & True & True & True \\ \hline
\end{tabular}
\end{center}

				\item If $P\Rightarrow Q$ and $Q\Rightarrow R$ are both true, then $P\Rightarrow R$ is true. \\
Proof: Looking at the proof directly above in rows a g e and h, which represent all combinations of P Q and R in which P $\Rightarrow$ Q and Q $\Rightarrow$ R are true, we can see that in column 6 that $P \Rightarrow R$ is always true.\\  
				\item $P\Rightarrow (Q\vee R)$ is equivalent to $\left(P\wedge (\sim Q)\right)\Rightarrow R$ and $\left(P\wedge (\sim R)\right)\Rightarrow Q$. \\
Proof: 
\begin{center}
\begin{tabular}{c||c|c|c|c|c|c|c||}
$$ & 1 & 2 & 3 & 4 & 5 & 6\\ \hline
$$ & P & Q & R & $P\Rightarrow (Q\vee R)$ & $\left(P\wedge (\sim Q)\right)\Rightarrow R$ & $\left(P\wedge (\sim R)\right)\Rightarrow Q$ \\ \hline
a & True & True & True & True & True & True \\ \hline 
b & True & True & False & True & True & True \\ \hline
c & True & False & True & True & True & True \\ \hline
d & True & False & False & False & False & False \\ \hline
e & False & True & True & True & True & True \\ \hline
f & False & True & False & True & True & True \\ \hline
g & False & False & True & True & True & True \\ \hline
h & False & False & False & False & False & False \\ \hline
\end{tabular}
\end{center}
We can see from the Truth table directly above that for all values in columns 4 5 and 6, all values are equal, therefore the three statements are equivalent.
				\item $P\Rightarrow (Q\wedge R)$ is equivalent to $\left(P\Rightarrow Q\right)\wedge \left(P\Rightarrow R\right)$. \\
Proof:
\begin{center}
\begin{tabular}{c||c|c|c|c|c||} 
$$ & 1 & 2 & 3 & 4 & 5 \\ \hline 
$$ & P & Q & R & $P\Rightarrow (Q\wedge R)$ & $\left(P\Rightarrow Q\right)\wedge \left(P\Rightarrow R\right)$ \\ \hline
a & True & True & True & True & True \\ \hline
b & True & True & False & False & False \\ \hline
c & True & False & True & False & False \\ \hline
d & True & False & False & False & False \\ \hline
e & False & True & True & True & True \\ \hline 
f & False & True & False & True & True \\ \hline
g & False & False & True & True & True \\ \hline
h & False & False & False & False & False \\ \hline
\end{tabular}
\end{center} 
Observing columns 4 and 5 of the truth table directly above, we can see that all the values are the same, therefore $P\Rightarrow (Q\wedge R)$ and $\left(P\Rightarrow Q\right)\wedge \left(P\Rightarrow R\right)$ are equivalent.
				\item If $P\Rightarrow R$ is true and $Q\Rightarrow R$ is true, then $\left(P\vee Q\right)\Rightarrow R$ is true. \\
Proof: First, we must realize that in order for wither statement to be false, R would have to be false. For the first statement $P \Rightarrow R$, P would have to be True, and in order for the second statement $Q \Rightarrow R$ to be False, Q would have to be True. Therefore, the only way both could be True while R is False is if P and Q were false, which we can see that for (P $\vee Q) \Rightarrow R$, the only way the statement could be False is if P or Q were True, which was shown to set former equations to false.
				\item If $\left(P\vee Q\right)\Rightarrow R$ is true, then $P\Rightarrow R$ is true. \\
Proof: According to truth table 1, $P \vee Q$ is true whenever P is true. Therefore, since the R remianed constant in the two propositions, it is therefore shown that when (P $\vee  Q) \Rightarrow Q$ is true, then so is P $\vee$ Q. 
			\end{enumerate}

			
\begin{enumerate}
	\item For statements (a)-(d), give a proof by just looking at the truth table for $\Rightarrow$, $\vee$, or $\wedge$.
		\vfill
	\item For statements (e) and (f), give a proof by making truth tables for the claimed equivalences.
	\vfill
	\newpage
	\item For statements (g) and (h), give a proof as follows:
		\begin{enumerate}
			\item Rewrite the $\Rightarrow$ in the formula $(P\vee Q)\Rightarrow R$ using $\vee$ and $\sim$, to get a formula that only involves $\vee$ and $\sim$.\vfill
			\item Do the same for $P\Rightarrow R$ and $Q\Rightarrow R$.\vfill
			\item After possibly doing some arithmetic, apply fact (c).
		\end{enumerate}
		\vfill
	\item The logical facts above will each prove to be useful in writing proofs. Below are a list of names or shorthands which correspond to logical facts (a)-(h). Match the logical fact to the name or shorthand. For some of these you will need to consult an outside source.
		\begin{enumerate}[(I)]
			\item {\em $\Rightarrow$ is transitive} d
			\item {\em specialization} c
			\item {\em prove each separately} f
			\item {\em modus tollens} a
			\item {\em a fortiori} h 
			\item {\em rule out the undesirable case} e
			\item {\em modus ponens / direct proof} b
			\item {\em proof by cases} g
		\end{enumerate}
\end{enumerate}
			
			
\end{document}
